Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-16 abstract/abstract.tex
Updates from Overleaf

Embracing a subtle novel approach, this process is facilitated by applying Siamese networks with a triplet loss, mimicking human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities. The study replaces traditional acoustical features with deep audio embeddings to compute high-level, sound-agnostic, and content-sensitive music identification.

@@

The preliminary results suggest that our approach to learning music-informed embeddings holds significant potential for nearly all MIR downstream tasks. Music-motivated embeddings represent a promising technique, adaptable potentially to other tasks hindered by data scarcity. This method could significantly influence advancements in intelligent music recommendation systems and the efficient enforcement of intellectual property rights.

>>>>

Embracing a subtle novel approach, this process is facilitated by using deep neural networks with a triplet loss, somewhat mimicking human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities. The study replaces traditional acoustical features with deep audio embeddings to compute high-level, sound-agnostic, and content-sensitive music identification.

@@

The preliminary results suggest that our approach to learning musically-informed embeddings holds significant potential for nearly all MIR downstream tasks. Music-motivated embeddings represent a promising technique, adaptable potentially to other tasks hindered by data scarcity. This method could significantly influence advancements in intelligent music recommendation systems and the efficient enforcement of intellectual property rights.

### 2023-05-16 figures/neural networks/Back_propagation.tex
Updates from Overleaf

\begin{SCfigure}[2\sidecaptionrelwidth][h]
	\centering
	\begin{tikzpicture}[shorten >=1pt]
      		\tikzstyle{unit}=[draw,shape=circle,minimum size =1.4cm]

@@

		\draw[->,red,line width=0.05cm] (2,-0.25) -- (0.75,0.3);
		\draw[->,red,line width=0.05cm] (2,2) -- (0.75,1.6);
    	\end{tikzpicture}
	\caption[Backpropagation of errors through the network.]{Once evaluated for all output units, the errors $\delta_i^{(L+1)}$ can be propagated backwards.}.
	\label{fig:error-backpropagation}
\end{SCfigure}

>>>>

\begin{SCfigure}[2\sidecaptionrelwidth][ht]
	\centering
	\begin{tikzpicture}[shorten >=1pt]
      		\tikzstyle{unit}=[draw,shape=circle,minimum size =1.4cm]

@@

		\draw[->,red,line width=0.05cm] (2,-0.25) -- (0.75,0.3);
		\draw[->,red,line width=0.05cm] (2,2) -- (0.75,1.6);
    	\end{tikzpicture}
	\caption[Backpropagation.]{\small{Backpropagation of errors through the network: once evaluated for all output units, the errors $\delta_i^{(L+1)}$ can be propagated backward.}}
	\label{fig:error-backpropagation}
\end{SCfigure}

### 2023-05-16 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time such as melody, harmony, rhythm, form, motif, counterpoint, etc. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

Abstract high-level features can be discerned regardless of waveform or production style. This approach provides an objective, comprehensive insight into a composition's musical content and meaning, much like analyzing sheet music uncovers its intentional foundation.

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint, so to say, derived from the propulsion of their melodic and harmonic contours through non-diatonic major thirds.

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}.

When applied to music, the principles of Gestalt psychology offer insights into how we perceive, organize, and interpret musical patterns. They suggest that our minds perceive auditory input much like visual input, seeking patterns and structures. This understanding significantly contributes to the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

>>>>

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time, such as melody, harmony, rhythm, form, motif, counterpoint, etc. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

Abstract high-level features can be discerned regardless of waveform or production style. This approach provides an objective, comprehensive insight into a composition's musical content, much like analyzing sheet music uncovers its behavioral foundation.

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint, so to say, derived from the envelope of their melodic and harmonic contours through non-diatonic major thirds.

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}. When applied to music, the principles of Gestalt psychology offer insights into how we perceive, organize, and interpret musical patterns. They suggest that our minds perceive auditory input much like visual input, seeking patterns and structures. This understanding significantly contributes to the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

### 2023-05-16 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage is not intended to deny the diversity and richness of other musical traditions worldwide but instead reflects the present study's focus on a specific cultural context.}, an essential component of human culture, history, and society, has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as it evolves. Studying music, however, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

@@

Music Information Retrieval (MIR) has emerged over the last few decades as an interdisciplinary field that combines engineering, musicology, and neuroscience. It aims to develop algorithms and techniques for automatically analyzing, classifying, and retrieving musical information, primarily from audio signals. While MIR research has traditionally focused on technical aspects such as Digital Signal Processing (DSP) and pattern recognition, it may have overlooked other relevant domains, including perception, cognition, music theory, aural skills, and musicology, along with their subfields.

>>>>

Music \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage is not intended to deny the diversity and richness of other musical traditions worldwide but instead reflects the present study's focus on a specific cultural context.}, an essential component of human culture, history, and society; it has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as it evolves. Studying music, therefore, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

@@

MIR has emerged over the last few decades as an interdisciplinary field that combines engineering, musicology, and neuroscience. It aims to develop algorithms and techniques for automatically analyzing, classifying, and retrieving musical information, primarily from audio signals. While MIR research has traditionally focused on technical aspects such as Digital Signal Processing (DSP) and pattern recognition, it may have overlooked other relevant domains, including perception, cognition, music theory, aural skills, musicology, and their subfields.

### 2023-05-16 introduction/Subsections/Motivation.tex
Updates from Overleaf

This research is conducted in collaboration with \textit{Epidemic Sound AB (ES)} \cite{EpidemicSite}, a prominent Swedish company that provides a global library of over 40,000+ royalty-free soundtrack music and 90,000+ sound effects, and the \textit{Music Technology Group (MTG)} at Universitat Pompeu Fabra in Barcelona, Spain.

The MTG specializes in developing new technologies related to sound and music, including music information retrieval, digital signal processing, computational models of music perception and cognition, and interactive music systems.

The collaboration with Epidemic Sound and the MTG aims to deepen our understanding of music and its fundamental structures, potentially leading to advancements and benefits for the company's technical products, services, and MIR academic research.

>>>>

This research is conducted in collaboration with \textit{Epidemic Sound AB (ES)} \cite{EpidemicSite}, and the \textit{Music Technology Group (MTG)} at Universitat Pompeu Fabra.

ES is a prominent Swedish company that provides a global library of over 40,000+ royalty-free soundtrack music and 90,000+ sound effects.

The MTG, based in Barcelona (Spain), specializes in developing new technologies related to sound and music, including music information retrieval, digital signal processing, computational models of music perception and cognition, and interactive music systems.

The collaboration with ES and the MTG aims to deepen our understanding of music and its fundamental structures, potentially leading to advancements and benefits for the company's technical products, services, and MIR academic research.

### 2023-05-16 introduction/Subsections/Objectives.tex
Updates from Overleaf

This study explores music's core tonal structure by constructing and analyzing an embedding space applicable to subsequent MIR tasks.

By blending MIR computational techniques and a musicological approach, we intend to examine a broad spectrum of musical works across diverse styles and genres to mimic human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities.

The research focuses on learning a single embedding, $e(M)$, for a music piece $M$ which encapsulates the underlying high-level content such as melody contour, harmony inter-relationships, and other holistic elements. The structuring of the embedding space will prioritize the clustering of high-level musical content while maintaining agnostic to sonic quality features.

>>>>

This study explores music's high-level tonal structure by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings applicable to subsequent MIR tasks.

The study aims to develop a unified embedding, $e(M)$, for a music piece $M$ that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components, arranging the embedding space to emphasize clustering of these sophisticated musical content without any consideration for sonic quality attributes. Utilizing a blend of Music Information Retrieval (MIR) computational methodologies and musicological perspectives, the research will explore musical compositions from various styles and genres.

This approach seeks to replicate human auditory capabilities in understanding and identifying abstract and semantic musical elements independent of their sonic qualities.

### 2023-05-16 introduction/Subsections/Previous work.tex
Updates from Overleaf

\item The resolution of the cross-modal text-to-music retrieval problem, enabling content creators to find music that matches the emotion conveyed in their text-based stories \cite{WonEmotionStories}.
\end{itemize}

Moreover, deep audio embeddings offer the advantage of transferability. Once trained, these embeddings can be utilized as a starting point for various other tasks, thus saving computational resources and time compared to training a model from scratch.

>>>>

\item The resolution of the cross-modal text-to-music retrieval problem, enabling content creators to find music that matches the emotion conveyed in their text-based stories \cite{WonEmotionStories}.

\item Music rearrangement is traditionally undertaken by expert music engineers or more classical MIR approaches \cite{Stoller2018IntuitiveTransitions}. This innovative approach presents a promising avenue for automated music rearrangement, enhancing the process's efficiency and quality. \cite{Plachouras2023MusicSegmentation}

\end{itemize}

Moreover, deep audio embeddings offer the advantage of transferability. Once trained, these latent representations can be utilized as a starting point for various other tasks, thus saving computational resources and time compared to training a model from scratch \cite{HamelTransferSimilarity}.

### 2023-05-16 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

This approach to learning is inspired by the way humans and animals learn from their environment, where a great deal of knowledge is acquired through observation and interaction without explicit instruction.

@@

Although there are resemblances between assimilation and accommodation in Piaget's theory and backpropagation in Deep Learning (DL), these processes occur in distinct mechanisms and contexts. Assimilation and accommodation modify mental structures to incorporate new information, while backpropagation adjusts neural network weights to enhance output predictions.

@@

\begin{definition}[Known invariance]
Known invariances are transformations or changes that the model or algorithm is explicitly designed to be robust against. This can include the input data's rotations, translations, scaling, or other known perturbations. By incorporating these invariances into the model's design, the algorithm is better equipped to generalize and recognize patterns in the presence of these transformations.
\end{definition}

@@

Unknown invariance: Unknown invariances are transformations the model has not been explicitly designed to handle. However, it still manages to be robust due to its learning capabilities. These invariances can emerge from the data and may not be known a priori. For example, a deep learning model might automatically learn to be invariant to certain lighting conditions or other factors not explicitly specified in its design.
\end{definition}

\begin{equation}
\label{eq:invariance_to_particular_transformation_function}
\forall x \in X, \forall y \in Y, \forall \rho \in \mathcal{T}, f(x) = f(y) \implies f(\rho(x)) = f(\rho(y))
\end{equation}

In expression \ref{eq:invariance_to_particular_transformation_function}, $X$ and $Y$ represent sets of input data, $\rho$ represents a transformation function, and $f$ represents a machine learning model. The expression states that for all inputs $x$ and $y$ in $X$ and $Y$, if $f(x) = f(y)$, then $f(\rho(x)) = f(\rho(y))$.

The expression specifies that a model is invariant to a particular transformation function $\rho$ if it produces the same output for both the original and transformed input.

@@

\item \textbf{Time-domain effects:} Time-domain manipulation alters the signal's timing or duration.
\begin{itemize}
    \item Speed change: Alters the playback speed of the signal by a factor $\alpha \in [0.9, 1.1]$.
    \item Stretch: Changes the signal duration without affecting its pitch by a factor $\beta \in [0.9, 1.1]$.
\end{itemize}

\item \textbf{Frequency-domain effects:} Frequency domain manipulation alters the frequency content or pitch of the signal.
\begin{itemize}
    \item Pitch shift: Changes the signal's pitch by $\Delta p \in [-1200, 1200]$ cents. Two-octave range, one octave higher to one octave lower.
\end{itemize}

\item \textbf{Nonlinear effects:} Nonlinear transformations to the audio signal, resulting in harmonic distortion or other complex changes.
\begin{itemize}
    \item Overdrive: Adds harmonic distortion to the signal based on the parameter $d \in [0, 30]$.
\end{itemize}

\item \textbf{Modulation effects:} Audio signal modulation using a low-frequency oscillator or control signal.
\begin{itemize}
    \item Chorus: Applies a varying time delay to the signal, resulting in a richer, thicker sound. The specified ranges determine the chorus parameters.
    \item Tremolo: Modulates the signal's amplitude at a modulation frequency $f_m \in [0.1, 100]$ Hz and depth $d_m \in [1, 101]$.
\end{itemize}

@@

\item \textbf{Noise effects:} These effects add noise or other random elements to the audio signal.
\begin{itemize}
    \item Additive noise: Adds a noise signal $N[n]$ with a specific signal-to-noise ratio (SNR) in the range $[12, 100]$.
\end{itemize}
\end{itemize}

>>>>

This approach to learning is inspired by the way humans and animals learn from their environment, where a great deal of knowledge is acquired through observation and interaction without explicit instruction. Self-supervised learning is a machine learning process where the model trains itself to learn one part of the input from another part of the input. In this process, the unsupervised problem is transformed into a supervised problem by auto-generating the labels.

@@

@@

\begin{definition}[Known invariance]
Known invariances are transformations or changes that the model or algorithm is explicitly designed to be robust against. By incorporating these invariances into the model's design, the algorithm is better equipped to generalize and recognize patterns in the presence of these transformations.
\end{definition}

@@

Unknown invariances are transformations the model has not been explicitly designed to handle. However, it still manages to be robust due to its learning capabilities. These invariances can emerge from the data and may not be known a priori. 
\end{definition}

@@

\item \textbf{Time-domain effects:} Alters the signal's timing or duration.
\begin{itemize}
    \item Speed change: Alters the playback speed of the signal by a factor $\alpha \in [0.9, 1.1]$.
    \item Stretch: Changes the signal duration without affecting its pitch by a factor $\beta \in [0.9, 1.1]$.
\end{itemize}

\item \textbf{Frequency-domain effects:} Alters the frequency content or pitch of the signal.
\begin{itemize}
    \item Pitch shift: Changes the signal's pitch by $\Delta p \in [-1200, 1200]$ cents. Two-octave range, one octave higher to one octave lower.
\end{itemize}

\item \textbf{Nonlinear effects:} Harmonic distortion or other complex changes.
\begin{itemize}
    \item Overdrive: Adds harmonic distortion to the signal based on the parameter $d \in [0, 30]$.
\end{itemize}

\item \textbf{Modulation effects:} Modulation using a low-frequency oscillator or control signal.
\begin{itemize}
    \item Chorus: Applies a varying time delay to the signal, resulting in a richer, thicker sound. The specified ranges determine the chorus parameters.
    \item Tremolo: Modulates the signal's amplitude at a modulation frequency $f_m \in [0.1, 100]$ Hz and depth $d_m \in [1, 101]$.
\end{itemize}

@@

\item \textbf{Noise effects:}
\begin{itemize}
    \item Additive noise: Adds a noise signal $N[n]$ with a specific signal-to-noise ratio (SNR) in the range $[12, 100]$.
\end{itemize}
\end{itemize}

### 2023-05-16 results/results.tex
Updates from Overleaf

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/PlotTimeSeriesResult}% 
\caption{This is an example of a figure and its caption.}
\label{fig:timeseries}
\end{figure}

\begin{table}[!ht]
\renewcommand{\arraystretch}{1.50}
\caption{This is an example of a table and its caption.}
\label{tablePCA}
\centering
\begin{tabular}{| c | c |}
\hline
\bfseries PCA & \bfseries Residual mean (in absolute values) \\
\hline\hline
Original PCA & 0.1267  \\
\hline
PCA on Centroid 1 & 0.1249\\
\hline
PCA on Centroid 2 & 0.1214  \\
\hline
\end{tabular}
\end{table}

>>>>

\input{figures/neural networks/manifold}

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/PlotTimeSeriesResult}% 
\caption[Graph.]{This is an example of a figure and its caption.}
\label{fig:timeseries}
\end{figure}

\begin{table}[!ht]
\renewcommand{\arraystretch}{1.50}
\caption[Table]{This is an example of a table and its caption.}
\label{tablePCA}
\centering
\begin{tabular}{| c | c | c | c | c |}
\hline
\bfseries PCA & \bfseries Residual mean & \bfseries XXX & \bfseries XXX & \bfseries XXX \\
\hline\hline
Original PCA & 0.1267 & XXX & XXX & XXX  \\
\hline
PCA on Centroid 1 & 0.1249 & XXX & XXX  & XXX\\
\hline
PCA on Centroid 2 & 0.1214  & XXX & XXX  & XXX\\
\hline
\end{tabular}
\end{table}
